# General Introduction {#gen-intro}

```{r 01-setup, include=FALSE}
library(plyr)
library(tidyverse)
library(lubridate)
library(readxl)
library(knitr)
library(kableExtra)
library(ggthemes)
library(plotly)
library(mousetrap)
library(randomizr)
library(data.table)
library(lavaan)
library(survival)
library(survminer)
library(papaja)
library(tidytidbits)

doc.type <- knitr::opts_knit$get('rmarkdown.pandoc.to')

################# DATA #################
GTA_Survival <- 
  read_rds("chapter_3_data/GTA_Survival.rds")

GTA_Survival2008 =
  GTA_Survival %>%
  filter(year(firstMLTrainingDate) > 2008 &
           (firstMLTrainingDate < firstMLAssessmentDate |
              is.na(firstMLAssessmentDate) == TRUE) &
           (SexId == 1 | SexId == 2)) %>%
  mutate(Gender = rename_factor(SexId, `1` = "Male", `2` = "Female"))

all_qual_annual_pathway_count <- 
  read_csv("../3 eshot/DLog_Analysis/SQL_Outputs/all_qual_annual_pathway_count.csv", na = "NULL")
distinct_qual_annual_pathway_count <- 
  read_csv("../3 eshot/DLog_Analysis/SQL_Outputs/distinct_qual_annual_pathway_count.csv", na = "NULL")
distinct_qual_annual_pathway_count$AwardId[distinct_qual_annual_pathway_count$AwardId== 1] <- 526
distinct_qual_annual_pathway_count$AwardId[distinct_qual_annual_pathway_count$AwardId== 2] <- 527
distinct_qual_annual_pathway_count$AwardId[distinct_qual_annual_pathway_count$AwardId== 3] <- 529
distinct_qual_annual_pathway_count$AwardId[distinct_qual_annual_pathway_count$AwardId== 5] <- 11
distinct_qual_annual_pathway_count$AwardId[distinct_qual_annual_pathway_count$AwardId== 18] <- 528

KEY_award <- 
  read_csv("../3 eshot/DLog_analysis/SQL_Outputs/KEYS/KEY_Award.csv")


if (is_latex_output()) {font = "serif" } else {font = "sans"}

```


## Mountain Training

Mountain Training is responsible for training walking, climbing, and mountaineering instructors in the United Kingdom and Ireland. Mountain Training is not one single entity; it is the outward-facing name for a group of organisations in the United Kingdom and Ireland. Mountain Training England, Cymru, Scotland, and the Mountain Training Board of Ireland are responsible for administering the Mountain Leader qualification in their respective countries. `r colorize("Whereas,")` Mountain Training United Kingdom and Ireland are responsible for the generic training pathway for all qualifications. However, as this project has stakeholders within each Mountain Training organisation, we simply refer to Mountain Training as a single entity throughout the thesis.

Mountain Training's qualifications all follow a similar pathway to qualification, which was created in 1964 for the Mountain Leadership Certificate (what is now the Mountain Leader qualification) and has not changed significantly since then. Candidates must first gain some *prerequisite experience and register for the qualification*, and then they *complete a training course*. Following the training course, candidates gain *further experience to consolidate skills*, and finally, they then need to *complete an assessment course*. Candidates are awarded the relevant qualification on successful completion of the assessment `r colorize ("course")`.

In 2018 there were `r printnum(all_qual_annual_pathway_count %>% filter(year == 2018) %>% select(pass_count), digits=0)` qualifications awarded to candidates, which suggests that this pathway is successful to some degree, as each year many candidates progress from registration to qualification. However, for all qualifications, the number of candidates at each stage of the pathway is lower than the number at the previous stage (i.e., registration to training, training to assessment, and passing an assessment). Figure \@ref(fig:qual-pathway-drop-off-plot) shows this drop-off for six of the Mountain Training qualifications. 


```{r qual-pathway-drop-off-plot, warning=FALSE, fig.height=5, fig.cap="Average number of candidates at each pathway stage 2009-2018.  LLA = Lowland Leader Award, CWI = Climbing Wall Instructor, ML = Mountain Leader - Summer, RCI = Rock Climbing Instructor, MLW = Mountain Leader - Winter, MCI = Mountaineering and Climbing Instructor."}
distinct_qual_annual_pathway_count %>%
  left_join(KEY_award %>% select(Id, Label, CategoryId),
            by = c("AwardId" = "Id")) %>%
  filter(between(year, 2009, 2018) &
           Label %in% c("LLA", "MLS", "MLW", "CWI", "RCI", "MCI")) %>%
  mutate(Label = factor(Label, c("LLA", "CWI", "MLS", "RCI", "MLW", "MCI")),
         Label = rename_factor(Label, "MLS" = "ML")) %>%
  group_by(Label) %>%
  summarise(
    Registered = mean(reg_count),
    Trained = mean(train_count),
    Assessed = mean(first_assess_count),
    Qualified = mean(first_time_pass_count)
    #reassess = mean(reassessment_count),
    #reassess_pass = mean(reassessment_pass_count)
  ) %>%
  melt(id = c("Label")) %>%
  ggplot(aes(x = variable, y = value, fill = variable)) +
  geom_col() +
  #geom_text(aes(label = value)) +
  guides(fill = FALSE) +
  facet_wrap(~ Label, ncol = 2, scales = "free_x") +
  labs(x = "Pathway stage", y = "Count") +
  coord_flip() +
  theme_few(base_size = 12, base_family = font) +
  theme(plot.margin = margin(l = 15, r = 5, t = 10, b = 5, unit = "pt"))
```


Mountain Training estimates that its qualification holders impact over 1.5 million people each year [@MTE2019]. This number is likely to increase in the coming years as the "The overall growth in numbers [of active adults] continue [*sic*] to be driven by strong upward trends in walking and adventure sports (a category which includes hill and mountain walking, climbing and orienteering)" [@SportEngland2020, p 14]. Therefore, it is vital to Mountain Training that they understand why people do and do not complete their qualifications.

This thesis focuses on the Mountain Leader qualification `r colorize("to understand")` the factors influencing the completion of Mountain Training qualifications, so that they can identify improvements that can be made to their training pathway. There are four main reasons for focusing on the Mountain Leader qualification: (a) it is the largest qualification as measured by the number of candidates; (b) it has the largest drop-off in candidates progressing from training to assessment, the drop-off at this point is of particular interest as candidates have engaged with the Mountain Training delivery system; (c) it is the highest entry-level qualification; and (d) it is the oldest qualification and has had few major changes `r colorize("made")` to it recently.


## The Mountain Leader Qualification

The Mountain Leader qualification is for "people who want to lead groups in the mountains, hills and moorlands of the UK and Ireland" [@MTUK2015a, p 5]. Whilst the Mountain Leader qualification is UK based and aimed at those who wish to lead others in the UK, training programmes in other countries are based on the success of the Mountain Leader qualification [@UIAA2015]. To qualify as a Mountain Leader, candidates must: (a) *register for the qualification* and gain a minimum of 20 Quality Mountain Days (QMDs),^[There is not a simple definition for a QMD; however, QMDs should "make a positive contribution towards a person’s development and maturity as an all round mountaineer" [@MountainTraining2019].] (b) complete a six-day *training course*, (c) gain a minimum of 20 additional QMDs as *further experience to consolidate skills (to give a minimum of 40 QMDs in total)*, and (d) successfully complete a five-day *assessment course*. Therefore, to become a Mountain Leader, `r colorize("a candidate")` must spend a minimum of 51 days in the mountains. Most successful candidates will have more experience than this, whether that is additional QMDs, other experience of mountain walking that does not meet the QMD criteria or other mountaineering experience. Therefore, becoming a Mountain Leader requires candidates to commit a significant amount of time and `r colorize("money.")`

Between 2009 and 2018 an average of 2,278 candidates registered for the Mountain Leader qualification each year, but only 559 qualified a year. When looking more closely at the numbers of candidates who did qualify, it becomes clear that there are two main components to qualifying: (a) getting to an assessment and (b) passing an assessment. Interestingly, most candidates did not get to an assessment (Figure \@ref(fig:survival-sex)), but most candidates who got to an assessment passed their first assessment (Figure \@ref(fig:PassRate)). It is also noteworthy that, as shown in Figure \@ref(fig:survival-sex), becoming a Mountain Leader is not a quick process (period between training and assessment, $M =$ `r GTA_Survival %>% filter(year(firstMLTrainingDate) > 2008 & Assessed == 1) %>% .$TimeToAssessment %>% mean() %>% printnum()` years, $SD =$ `r GTA_Survival %>% filter(year(firstMLTrainingDate) > 2008 & Assessed == 1) %>%  .$TimeToAssessment %>% sd() %>% printnum()`).


To examine the difference in the number of candidates being trained and assessed for the Mountain Leader qualification in more detail we carried out a *survival analysis* [cf. @HarrellJr2015]. With this survival analysis, rather than looking at summary statistics averaged over several years, we look at the probability of an individual candidate having been assessed over time following their training course. As can be seen in Figure \@ref(fig:survival-sex), at any given point in time, fewer female candidates get to an assessment than male candidates. The percentage likelihood of a candidate having been assessed five years following their training course is ~32% and ~40% respectively for female and male candidates. After this point the rate of candidates being assessed decreases for both genders. Over half of candidates who did reach assessment did that within 18 months of their training courses, but it `r colorize("was")` not unusual to take longer, and some candidates `r colorize("were")` assessed more than five years after their training course.


```{r}
cap <- 
  paste0("Survival rates for female and male candidates post-training. Candidates trained 2009-2019 ($N = ", 
         printnum(as.numeric(filter(GTA_Survival, year(firstMLTrainingDate) > 2008) %>% nrow()), digits = 0), 
         "$). The green dotted and dashed line represents the median time to assessment, and the blue dashed line represents the mean time to assessment.")
```


```{r survival-sex, warning=FALSE, fig.cap=cap, fig.height=3}
GTA_Survival2008 =
  GTA_Survival %>%
  filter(year(firstMLTrainingDate) > 2008 &
           (firstMLTrainingDate < firstMLAssessmentDate |
              is.na(firstMLAssessmentDate) == TRUE) &
           (SexId == 1 | SexId == 2)) %>%
  mutate(Gender = rename_factor(SexId, `1` = "Male", `2` = "Female"))

surv_object <-
  Surv(time = GTA_Survival2008$TimeToAssessment,
       event = GTA_Survival2008$Assessed)

fit1 <-
  survfit(surv_object ~ Gender, GTA_Survival2008)

ggsurvplot(
  fit1, 
  # data = GTA_Survival2008,
  fun = "event",
  surv.scale = "percent",
  pval = FALSE,
  # pval.coord = c(4, .05),
  # xlim = c(0, 5),
  # ylim = c(0,.45),
  break.x.by = 1,
  break.y.by = .05,
  xlab = "Time after training in years",
  ylab = "% of candidates assessed",
  censor = FALSE,
  conf.int = TRUE,
  ggtheme = theme_few(base_size = 12,
                      base_family = font),
  tables.theme = 
    theme_survminer(font.main = 10, font.x = 10,
                    font.y = 10, base_family = font,
                    base_size = 10)
) -> p

p$plot + 
  theme(legend.position = c(.7,.1), legend.direction = "horizontal") +
  rremove("legend.title") +
  geom_vline(xintercept = (GTA_Survival2008 %>% filter(Assessed == 1) %>%  .$TimeToAssessment %>% mean()), lty = 2, colour = "blue") +
  geom_vline(xintercept = (GTA_Survival2008 %>% filter(Assessed == 1) %>%  .$TimeToAssessment %>% median()), lty = 4, colour = "green3")
```


We also examined the pass rates for the Mountain Leader qualification. The pass rate `r colorize("increased")` over time, and there `r colorize("were")` changes in pass rates over the last 10 years for female and male candidates (Figure \@ref(fig:PassRate)). When looking at pass rates for the last 10 years, female candidates are less likely to pass their first assessment, but the pass rate `r colorize("increased")` faster for them than it `r colorize("did")` for male candidates. However, when looking at data from the last five years, neither the effect of gender on the pass rate or rate of change of the pass rate is statistically significant.


```{r}
cap <- paste0("Pass rates for female and male candidates assessed since 2000 ", "($N = ", printnum(as.numeric(nrow(GTA_Survival %>%
  filter(Assessed == 1 & year(firstMLAssessmentDate) > 1999 & 
           (SexId == 1 | SexId == 2)))), digits = 0), "$).")
```


```{r PassRate, fig.cap=cap, fig.height=3}
GTA_Survival %>%
  filter(Assessed == 1 & year(firstMLAssessmentDate) > 1999 & 
           (SexId == 1 | SexId == 2)) %>% 
  mutate(SexId = rename_factor(SexId, `1` = "male", `2` = "female")) -> df

df %>%
  mutate(
    Gender =
      rename_factor(
        SexId,
        "male" = 
          paste("Male \nn =",
                printnum(
                  as.numeric(nrow(filter(df, SexId == "male"))), digits = 0)),
        "female" = 
          paste("Female \nn =",
                printnum(
                  as.numeric(nrow(filter(df, SexId == "female"))), digits = 0))
      ),
    firstMLAssessmentDate =
      round_date(firstMLAssessmentDate, unit = "week")
  ) %>%
  ggplot(aes(x = firstMLAssessmentDate, 
             y = 100 * FirstTimePass, 
             colour = Gender, fill = Gender)) +
  geom_smooth(method = "loess") +
  coord_cartesian(xlim = c(ymd(20100101), ymd(20200101))) +
  labs(x = "Assessment Year",
       y = "First Time Pass Rate %") +
  theme_few(base_size = 12, base_family = font) +
  # theme(legend.position = "bottom")
  theme(legend.position = c(.5, .85), legend.direction = "horizontal")
```


## Thesis Rationale

Mountain Training wanted to understand why more people do not complete the Mountain Leader qualification and identify if there `r colorize("are")` any changes that they could make to the pathway that would better support their candidates. It is unlikely that there is a single factor that would be a “silver bullet” to improve completion rates. Instead, there is likely a myriad of factors which influence completion at various stages of the pathway. Some of these factors will be generic to all candidates, whilst some may be specific to individual `r colorize("and/or groups of")` candidates.

The work reported in this thesis is the first empirical investigation of the factors that influence the completion rate for the Mountain Leader qualification. Therefore, this thesis is an essential first step for Mountain Training towards making evidence-based changes to their training pathway to improve the completion rates of their qualifications.


## Terminology 

It is important that I provide a note on the terminology used in this thesis. Historically, and somewhat incorrectly, the terms *sex* and *gender* have been used somewhat interchangeably in the scientific literature. Current guidelines from the @APA2020 define gender as, "the attitudes, feelings, and behaviors that a given culture associates with a person’s biological sex" [@APA2012, p 12]; sex as, "biological sex assignment;" and \textit{gender identity} as "a component of gender that describes a person’s psychological sense of their gender" [@APA2020, Section 5.5].

Based on the definitions of sex and gender presented above, it would be more appropriate to use terms such as "man" and "woman" than "female" and "male" when discussing gender differences. However, in the studies reported in this thesis, I retrieved data concerning participants' gender from Mountain Training's Candidate Management System (CMS), which stores it as female, gender-neutral, and male. To avoid making presumptions about participants gender identities, we have used the terms female and male throughout this thesis. No data were collected from gender-neutral candidates in any of the studies presented in this thesis. This may be unsurprising, given that 0.03% of all Mountain Training candidates report being gender-neutral [@MountainTrainingUK2019a]. Further, following the same principle, when discussing previous research, we have used terminology consistent with that used by the original authors.


## Pathways to Expertise

The development of expertise and the pathways to achieving it are of interest in a variety of domains. Historically, most studies have examined the impact of specific factors on the completion of a training pathway [e.g., delays in completing PhDs; @VandeSchoot2013a]. However, these single-variable approaches fail to acknowledge that there may be vast differences between individuals in their pathways to expertise.

Developing expertise is the result of complex interactions between a variety of developmental factors [e.g., practice and training, personality traits, motivation, social support to name but a few; @Baker2013; @Gagne2004; @Johnston2018]. Further, different factors will be more salient at different points of development pathways [@Rees2016]. As such, a growing number of researchers now recommend that multidisciplinary approaches should be adopted for identifying the complex interactions that influence talent/expertise development [e.g., @Abernethy2013; @Gullich2019; @Johnston2018; @Pearson2006; @Rees2016; @Schorer2013]. Based on these principles, several projects in the elite sport domain have recently used multidisciplinary studies to explore the most important combinations of factors in the development of athletes in elite pathways [e.g., @Gullich2019; @Hardy2017; @Jones2019a; @Jones2019b; @Jones2020; @Rees2016]. These projects have used a mixed-methods approach, collecting both rich qualitative data from athletes and making use of state-of-the-art machine learning techniques to identify sets of variables, whose main-effects and interactive-effects are able to discriminate between athletes at different performance levels.

In line with the studies mentioned above, in this thesis, I took a holistic approach in understanding the factors that influence the completion rate of the Mountain Leader qualification, where we explored many factors using a variety of analytical methods. The work in the present thesis considers a wide range of factors. Providing a detailed literature review of `r colorize("factor")` each is beyond the scope of the present thesis, given the number of constructs included. However, in the following section, I aim to provide a brief overview of relevant theory and explain its relevance to the Mountain Leader qualification. Empirical chapters include additional detail where relevant. Further, these theories provide useful frameworks for discussing the results of the empirical chapters. These theories are useful for understanding the relationships between various factors, but this thesis is not a test of any specific theory.


## Personality and Individual Differences {#gen-intro-paid}

It is widely accepted that people are different from one another and that the characteristics that make them individuals influence thoughts, feelings, and behaviours [cf. @Roberts2015]. The Big Five model of personality [@McCrae1987] is well established in the psychology literature and has been widely used when considering individual differences in a wide range of outcomes [e.g., behaviour-change, performance, motivation; cf. @Allen2013; @Ryan2019]. Given the pervasive influence of personality, and specifically the Big Five, on human functioning, it seems reasonable to expect that some aspects of the Big Five might have relevance for the Mountain Leader qualification. For example, more conscientious individuals will be more hardworking and ambitious than those who are less conscientious [@McCrae1987] and will persevere through difficulty. Consequently, conscientious individuals may be more likely to complete the Mountain Leader qualification than those who are less conscientious.  As another relevant example, extraversion has been positively associated with effective leadership and decision making [@Hardy1996; @Judge2002]. Personality constructs beyond the Big Five are also likely to be relevant. For example, higher levels of resilience are associated with a range of positive outcomes, including overcoming adversity [cf. @Smith2008]. Becoming a Mountain Leader is a long process and requires candidates to operate in challenging environmental conditions (e.g., wind and rain), in addition, they are likely to encounter other obstacles in life. Therefore, the more resilient a candidate is, the more likely they are to overcome such adversity and obstacles.


## Motivation {#gen-intro-sdt}

`r colorize("Candidates engage with the Mountain Leader qualification for different reasons. Therefore, in trying to understand why some candidates complete the Mountain Leader qualification and others do not, it is important to consider these reasons and the associated motivation. Further, it is widely accepted that a raft of factors will influence motivation")` [@Deci2000; @Weinberg2014], `r colorize("these factors include both individual characteristics (e.g., personality, individual goals) and situational factors (e.g., interaction with others, the context of the behaviour); understanding candidates’ individual characteristics and situations is, therefore, important when trying to understand the effects of motivation on candidates likelihood of completing the Mountain Leader qualification.")`

`r colorize("In this thesis I use self-determination theory")` [@Deci1985b; @Deci2000; @Ryan2017] `r colorize("to consider motivation its relationship with both candidates' individual characteristics and situational factors. Self-determination theory is a theory of human behaviour that is widely used when studying motivation")` [cf. @Ryan2019] `r colorize("and grew from studies of")` *intrinsic motivation* [e.g., @Deci1975; @Deci1980a]. Intrinsically motivated behaviours are "behaviours that are motivated by the underlying need for competence and self-determination .... that are performed in the absence of any apparent external contingency" [@Deci1980a, p 42].

Self-determination theory contains six mini-theories: *cognitive evaluation theory* [@Deci1975; @Deci1980], *organismic integration theory* [@Deci1985b; @Ryan1989], *causality orientations theory* [@Deci1985a], *basic psychological needs theory* [@Ryan2000a], *goal contents theory* [@Kasser1996; @Niemiec2009], and *relationship motivation theory* [@Deci2014; @Ryan2017]. Basic psychological needs theory and organismic integration theory are particularly relevant to the present study and are described below.


### Basic Psychological Needs Theory.

*Basic psychological needs theory* is central to self-determination theory and each mini-theory [@Ryan2002]. Basic psychological needs theory posits three basic psychological needs: *autonomy*, feelings of volition, choice, and internal control; *competence*, the feeling of mastery and effectiveness; and *relatedness*, feeling connected and involved with others. Considerable evidence supports self-determination theorists' suggestion that need satisfaction is essential for optimal-functioning, good mental health, and well-being  [e.g., @Chen2015; @Deci2000]. Further, self-determination theorists suggest that the satisfaction of `r colorize("these three needs")` supports well-being and "high-quality" motivation, as the satisfaction of these needs will lead people to invest more in the activities or goals they are actively pursuing, whereas the frustration of these needs will lead people to invest less in `r colorize("their activities or goals")` [@Ryan2019]. Accordingly, it seems reasonable to suggest that candidates whose basic psychological needs are satisfied rather than frustrated in the pursuit of becoming a Mountain Leader will invest more effort and will therefore be more likely to complete the qualification.


### Organismic Integration Theory.

*Organismic integration theory* is concerned with extrinsic motivation and activities where the outcome is separable from the behaviour [@Ryan1985]. Self-determination theory has traditionally suggested that every behaviour can be placed on a continuum, the relative autonomy continuum,  with autonomous or self-determined motives on one side and controlled or non-self-determined motives on the other, and that people will therefore vary in both levels of and quality of motivation [@Deci1985b; @Sheldon2019].^[There is some evidence that the various types of motivation should not be placed on a continuum but may be better considered as contiguous [@Chemolli2014]. That notwithstanding, it is still considered that the different types of motivation may `r colorize("engender")` different outcomes.]

Organismic integration theory suggests that there are four types of extrinsic motivation---despite suggesting that they form a continuum. Two of these are considered controlled forms of extrinsic motive and the other two are considered autonomous forms of extrinsic motive. The most controlled form of extrinsic motivation `r colorize("that is")` proposed is *external regulation*, where an individual is motivated by external pressures or reward. Whist external regulation can be a powerful form of motivation, it is not typically thought of as having a long-lasting influence on behaviour [@Ryan2000c]. The next form of extrinsic motivation is *introjected regulation*, where behaviour is motivated by internally controlling pressures (e.g., ego-involvement, contingent self-worth). This form of motivation may also be powerful but may be weakened in the face of setbacks and behaviours that result from introjected regulation are unlikely to be stable [@Deci1995; @Ryan2019].

Considering the more autonomous forms of extrinsic motivation, *identified regulation* is when an individual consciously values their engagement in the activity and accepts the behaviour as personally important. Finally, the most autonomous form of extrinsic motivation proposed is *integrated regulation*, when the behaviour is congruent with personally endorsed values. Integrated regulation shares some features with intrinsic motivation; however, it is considered an extrinsic form of motivation as the outcome is separable from the behaviour [@Ryan2002]. 

Self-determination theory suggests that autonomous forms of motivation are more sustainable and better predictors of performance and goal persistence than controlled forms of motivation are [@Hagger2015; @Pelletier2001; @Ryan2019]. Various studies have found evidence that individuals who engage in behaviours for autonomous reasons will experience more positive affect, feel more satisfied, persist for longer, and feel more competent [cf. @Frederick-Recascino2002]. Therefore, it is likely that candidates with more autonomous forms of motivation will be more likely to become Mountain Leaders than those with more controlled forms of motivation, especially as time goes on.


### Hierarchical Models of Motivation.

Researchers have suggested that motivation is a complex construct and operates in a hierarchical fashion on at least three different levels [e.g., @Ingledew2009; @Vallerand1997; @Vallerand2002]. These levels are the global/dispositional, (why an individual generally engages in activities), contextual/participatory (the contents of motives within a particular domain or what an individual is trying to achieve or avoid), and situational/regulatory motives [the perceived locus of causality of the behavioural goals---where the motive sits on the relative autonomy continuum; @Deci2000; @Ingledew2009; @Vallerand1997; @Vallerand2002]. At the global level, motivation is considered as an individual difference, and therefore results in general consequences. Whereas, at the participatory level, the context will influence motivation and motivation may be manipulated more easily [@Vallerand2002]. Research has shown that these different levels of motivation have a top-down influence; for example, participatory motives influencing regulatory motives [@Ingledew2009]. Given the evidence for`r colorize(", and the effects of,")` these different levels of motivation, it may be important to consider the multiple levels of motivation for becoming a Mountain Leader in the present project.


## Self-Efficacy Theory {#gen-intro-self-efficacy}

Self-efficacy refers to an individual's confidence in their ability to carry out a specific task at a given time. Self-efficacy theory suggests that if an individual possesses the necessary skills and is sufficiently motivated, then their level of self-efficacy will be the primary determinant of their performance, how much effort they will put in, and how long they will persist---particularly in the face of adversity [@Bandura1977; @Bandura1982; @Bandura1997]. @Bandura1982 suggested that four main sources of information determine an individual's perception of self-efficacy.^[Other researchers [e.g., @Maddux1995] have added additional sources; however, most research considers the four sources proposed by Bandura.] In decreasing order of influence on efficacy beliefs, they are *previous performance accomplishments*, *vicarious experience/modelling*, *social/verbal persuasion*, and *physiological/emotional states* [cf. @Bandura1982; @Samson2011]. 

Results of meta-analyses offer evidence supporting the relationships suggested by Bandura between self-efficacy and performance and persistence across several domains, for example in sports [@Moritz2000], in work contexts [@Stajkovic1998], and academic environments [@Multon1991]. Several experimental lab studies `r colorize("that manipulated")` participants' level of self-efficacy found that higher levels of self-efficacy are related to increased task persistence [@Hutchinson2008; @Tenenbaum2001; @Weinberg1979; @Weinberg1980; @Weinberg1981]. Self-efficacy theory, and the evidence supporting it, suggests that candidates who have higher levels of self-efficacy are more likely to become Mountain Leaders.


## Theory of Planned Behaviour {#gen-intro-social-tpb}

The theory of planned behaviour [@Ajzen1991; @Ajzen1986] is a model for predicting behaviour and builds on the theory of reasoned action [@Ajzen1980] and that intentions are the best predictor of future behaviour. The theory suggests that an individual's intentions are the result of their attitudes towards the behaviours and the *subjective norm* surrounding that behaviour. Subjective norms are formed based on an individual's belief about other's views and the motivation of the individual to comply with those views. 

The theory of planned behaviour extends the theory of reasoned action by including *perceived behavioural control* as an influence on both the intention and the behaviour. An individual's perceived behavioural control reflects the extent to which they feel that they have the resources and opportunities to control the behaviour [n.b., @Ajzen1991 suggested that perceived behavioural control is similar to the construct of self-efficacy]. Meta-analytic studies provide evidence that the constructs that comprise the theory of planned behaviour can predict future behaviour [@Armitage2001; @Hagger2002]. 

It should be noted that the theory of planned behaviour is not without its critics, some of whom suggest that the theory is too limited in the constructs that it includes [cf. @Sniehotta2014]. However, the present project considers more constructs than just those related to the theory of planned behaviour, thus allaying such concerns. The theory of planned behaviour suggests that candidates who feel that it is normal to complete things in general, or more specifically the Mountain Leader pathway, and intend to complete the pathway will be more likely to complete than those who do not. Further candidates with higher rather than lower levels of perceived behavioural control will be more likely to become Mountain Leaders because they feel that they have the resources and opportunities to control the behaviour.


## Support {#gen-intro-support}

The literature introduced above relates primarily to candidates. However, it is clear that candidates will be influenced by `r colorize("other people")` during the journey through the Mountain Leader qualification pathway. Importantly to Mountain Training, the course staff will influence candidates; indeed, course staff could be thought of as `r colorize("coaches to")` candidates. In the sections that follow, I make the distinction between coaching and social support, where coaching relates to specific behaviours that the course staff may or may not engage in [@Wagstaff2018] and social support is the support that candidates may or may not receive in a broader context [@Cutrona1990a].


### Coaching. {#gen-intro-social-coaching}

Coaching usually aims to improve an individual's knowledge, skills, and competencies [@Wagstaff2018]. Leaders may engage in coaching behaviours, and some models of leadership [e.g., transformational leadership theory; @Bass1985] include elements of coaching behaviours. Mountain Leader course staff will engage in coaching behaviours to a greater or lesser extent, and their facilitation of candidate's development will vary accordingly. @Wagstaff2018 describe five coaching behaviours, based on sport and business coaching models: (a) observing and performance analysis, (b) asking effective questions, (c) facilitating goal setting, (d) providing developmental feedback, and (e) providing motivational feedback. 

In addition, coaches' actions may support candidates' basic psychological needs to a greater or lesser extent by tailoring the course delivery to individual candidates on the course [@Arthur2019; @Markland2010]. Supporting basic psychological needs is relevant to the present project as organismic integration theory suggests that social factors that support the basic psychological needs foster the development of more autonomous forms of extrinsic motivation [@Deci1991; @Ryan1995; @Ryan2002]. Autonomy support, the provision of structure, and involvement are need supportive behaviours, which support the development of autonomous regulation [@Ryan2016; @Markland2010; @Silva2010]. Therefore, it seems reasonable to expect that candidates who felt their course staff engaged in more coaching behaviours and did so in a *need supportive* manner are more likely to become Mountain Leaders than candidates who did not feel that their course staff did.


### Social Support. {#gen-intro-social-support}

Social and sports psychology widely consider four types of social support: *esteem support*, bolstering a person's sense of competence or self-esteem (e.g., giving individuals positive feedback on their skills and abilities, expressing a belief that the person is capable of coping with a stressful event); *emotional support*, the ability to turn to others for comfort and security during times of stress, leading the person to feel that he or she is cared for by others; *informational support*, providing the individual with advice or guidance; and *tangible support*, concrete assistance, providing someone with the necessary resources (e.g., financial assistance, physical help with tasks) to cope with something [e.g., @Cutrona1990a; @Cutrona1990b; @Freeman2010; @Freeman2014; @Rees2007a; @Rees2007]. 

Social support has been associated with increased levels of self-confidence, has "stress-buffering effects," and can be used to influence self-efficacy [@Rees2007; @Freeman2014; @Samson2011]. However, the effects of perceived support (i.e., the support an individual feel they would have available to them should they need it) are not the same as the effects of received support [the frequency with which support that has been provided to an individual; cf. @Rees2007; @Freeman2014]. Therefore, it is likely that candidates who feel they have more social support available to them *and* those who receive more social support are more likely to become Mountain Leaders than those who feel that their needs are not met. Distinguishing between the effects of perceived and received support will be important for understanding the implications of any social support related findings.


## Thesis Structure

A substantial body of work is reported in appendices, and the empirical chapters have been written in a way that aids their preparation for publication. The reasons for this are twofold. Firstly, it was necessary to develop the data collection tool needed for the work reported in Chapter \@ref(ml-pra). This work was foundational in the PhD process, in both developing the researcher and enabling the completion of five of the six main studies. However, including it in the main body of the thesis `r colorize("may")` distract the reader and likely detract from the "story" presented in the thesis. Therefore, this development work is reported in Appendix \@ref(survey-tool-dev) so that the reader can engage with it having read the main body of the thesis, which reports on the work answering the research question/s. Secondly, this thesis aims to satisfy the dual objective of writing a thesis and preparing the research for publication.

A consequence of writing this thesis with publication in mind is the self-referential terminology used in the empirical chapters. Accordingly, empirical chapters are written in the first-person plural, consistent with conventions in reporting co-authored research, and the remaining chapters are written in the first-person singular. The format of the thesis is in line with the policy of the School of Sport, Health and Exercise Sciences.

There are three empirical chapters following this chapter, which report the results of six studies. Chapter \@ref(ml-qualitative) presents a qualitative inquiry that aimed to explore organisational managers' understanding of factors that influence the completion of the Mountain Leader qualification (Study 1). Chapter \@ref(ml-pra) comprises three studies, each of which uses state of the art pattern recognition techniques to identify key discriminatory features that provide insight into the factors influencing different stages of the qualification pathway (Studies 2-4). Chapter \@ref(self-efficacy-qmds) explores the main and interactive effects of gender and experience on self-efficacy; in this chapter, I present two studies (Studies 5 and 6). Study 5 presents the development and validation of a measure for Mountain Leader related self-efficacy. Study 6 presents the findings of a study examining relationships between gender and experience on self-efficacy. Chapter \@ref(general-discussion) is a discussion of the three empirical chapters, identifying consistent themes and their implications as well as identifying potential future directions for research. Chapter \@ref(impact-activities) then describes the dissemination of this research to the Mountain Training network. Finally, the appendices provide supplementary information, development work, additional analyses, and a summary of other work carried out during the PhD that has been relevant to my development as a researcher.