\frontmatter

\section*{Declaration and Consent}
\addcontentsline{toc}{section}{Declaration and Consent}

Yr wyf drwy hyn yn datgan mai canlyniad fy ymchwil fy hun yw’r thesis hwn, ac eithrio lle nodir yn wahanol. Caiff ffynonellau eraill eu cydnabod gan droednodiadau yn rhoi cyfeiriadau eglur. Nid yw sylwedd y gwaith hwn wedi cael ei dderbyn o’r blaen ar gyfer unrhyw radd, ac nid yw’n cael ei gyflwyno ar yr un pryd mewn ymgeisiaeth am unrhyw radd oni bai ei fod, fel y cytunwyd gan y Brifysgol, am gymwysterau deuol cymeradwy.

I hereby declare that this thesis is the results of my own investigations, except where otherwise stated.  All other sources are acknowledged by bibliographic references. This work has not previously been accepted in substance for any degree and is not being concurrently submitted in candidature for any degree unless, as agreed by the University, for approved dual awards.


\section*{Funding}
\addcontentsline{toc}{section}{Funding}

This PhD was funded by a Knowledge Economy Skills Scholarship (KESS 2), which was part funded by Mountain Training UKI and by the Welsh Government's European Social Fund.

Mae’r Ysgoloriaeth Sgiliau Economi Gwybodaeth (KESS 2) yn fentersgiliau lefel uwch Cymru gyfana arweinir gan Brifysgol Bangor ar ran y sector AU yng Nghymru. Fe’i cyllidir yn rhannol gan raglen cydgyfeirio Cronfa Gymdeithasol Ewropeaidd (ESF)ar gyfer Gorllewin Cymru a’r Cymoedd.

Knowledge Economy Skills Scholarships (KESS 2) is a pan-Wales higher level skills initiative led by Bangor University on behalf of the HE sector in Wales. It is part funded by the Welsh Government’s European Social Fund (ESF) convergence programme for West Wales and the Valleys.

\newpage

\section*{Acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}

Lew, thank you for taking on a climber who wanted to know more and affording me ``learnings" through your supervision. I can't express how important your support and guidance has been, academic and beyond.

To my family, you have no idea how important you have been in getting me to where I am now. 

To my friends, old and new, thank you all for making sure that there has always been someone there for me; from days (or even weeks) in ``the nature" to having conversations about statistical minutiae over coffee, your friendships have all been invaluable to me. Chris, you will never know the influence that you have had on my life, but thank you.

Ross, you have provided me with an incredible level of support---more than anyone could ask for. I am certain that I would not be in the position I am in today was it not for your support. Thank you for guiding me through this journey.

